# Import Data

Collection of syntax snippets demonstrating how to import data into R.

Import from:

- Text File
- Excel File
- Statistics Package
- Database

## Text File

There are two different ways to import text files.

- Base R
- Hmisc

### Base R

```{r CsvExampleBase eval=FALSE, echo=TRUE}
foo <- read.csv(file = FileName,
                ,header = TRUE
                ,sep = ","
                ,quote = "\""
                ,dec = "."
                ,fill = TRUE
                ,comment.char = "#"
                )
```

### Hmisc

_Import a Single File_

```{r CsvExampleHmisc eval=FALSE, echo=TRUE}
library(Hmisc)
file_name <- "data/file.csv"
foo <- csv.get(file = FileName
               ,datevars = c()
               ,dateformat = "%m/%d/%Y"
               ,comment.char = "#"
              )
```

_Import Multiple Files_

This imports a series of files and concatenates them together.

```{r CsvExampleMultiple eval=FALSE, echo=TRUE}
library(Hmisc)
file_list <- list.files(pattern="*.csv")

for (file in file_list)
{
  # if the merged dataset doesn't exist, creates it 
  if (!exists("tbl"))
    {
      tbl <- csv.get(file = file
                     ,datevars = c()
                     ,dateformat = "%m/%d/%Y"
                     ,comment.char = "#"
                     )
    }
  else
    {
      tmp <- csv.get(file = file
                     ,datevars = c()
                     ,dateformat = "%m/%d/%Y"
                     ,comment.char = "#"
                     )
      tbl<-rbind(tbl, tmp)
      rm(tmp)
    } 
}
```


## Excel File

```{r ExcelExample eval=FALSE, echo=TRUE}
library(XLConnect)
file_name <- "data/foo.xlsx"
workbook <- loadWorkbook(file_name, create = FALSE)
readWorksheet(
              object = workbook
              ,sheet = 'Sheet1'
              ,region = 'A1:Z10'
              ,header = TRUE
              ,forceConversion = FALSE
              ,dateTimeFormat  # Defaults to getOption("XLConnect.dateTimeFormat")
              ,check.names=TRUE
             )
```


## Statistics Packages

### SPSS File

```{r SpssExample eval=FALSE, echo=TRUE}
library(foreign)
file_name <- "data/foo.bar"
foo <- read.spss(file = file_name,
                 use.value.labels = TRUE,
                 to.data.frame = FALSE,
                 max.value.labels = Inf,
                 trim.factor.names = FALSE,
                 trim_values = TRUE,
                 reencode = NA,
                 use.missings = to.data.frame
                )
```

### Stata DTA file

```{r StataExample eval=FALSE, echo=TRUE}
library(foreign)
file_name <- "data/foo.bar"
foo <- read.dta(file = file_name,
                convert.dates = TRUE,
                convert.factors = TRUE,
                missing.type = FALSE,
                convert.underscore = FALSE,
                warn.missing.labels = TRUE
               )
```


## Database

```{r RodbcExample eval=FALSE, echo=TRUE}
library(RODBC)
con <- odbcConnect("DSN", "UserID", "PassWord")
qry <- paste(
             'SELECT * '
             ,'FROM tblName'
             ,'WHERE foo=foo'
             ,'ORDER foo'
             ,sep=" "
            )
data_frame <- sqlQuery(con, qry)
```

If you are using the Analytics Template system, you should configure
the "odbcConnect()" string in config.R file.

```{r RodbcExample eval=FALSE, echo=TRUE}
con <- odbcConnect("DSN", "UserID", "PassWord")
```

Another nice way to read in a query from an external file is this:

```{r RodbcExample eval=FALSE, echo=TRUE}
qry <- readLines("sql/query_file.sql")
data_frame <- sqlQuery(con, qry)
```
